Revisiting Short-Chain Fatty Acids and the Microbiota in Colorectal Cancer

### 2018-05-24 submission/manuscript.Rmd
add abstract template

>>>>

**Background.** Certain bacterial community members have been shown to increase colon tumorigenesis while other taxa have been associated with normal colons. The some of the taxa associated with normal colons can use fiber to produce short-chain fatty acids which have been shown to inhibit tumor growth in model systems. However, the data supporting the importance of SCFAs in humans is less certain. Here we test the hypothesis that short-chain fatty acid concentration decreases in individuals with colorectal tumors.

**Methods.** We analyzed a cross-sectional (n=`r length(metai$sample)`) and pre- and post-treatment (n=`r length(metaf$initial)`) group for their concentrations of acetate, butyrate, and propionate. Analysis also included tumor classification models using Random Forest, imputed gene relative abundance with PICRUSt, and metagenomic sequencing on a subset (n=`r length(geof_samples$X30)`) of the total cross-sectional group.

**Results.** No difference in short-chain fatty acid concentration was found between normal individuals and patients with adenoma or carcinoma (P-value `r ifelse(min(pull(hplc_groups_scfa_kruskal, bh)) > 0.15, invisible(">"), invisible("<"))` 0.15). There was no difference in the ability of classification models with or without short-chain fatty acids added to them in predicting adenoma or carcinoma (P-value `r ifelse(min(pull(disease_rf, bh)) > 0.05, invisible(">"), invisible("<"))` 0.05). Using metagenomic sequencing, there was also no difference in genes involved with short-chain fatty acid synthesis (P-value `r ifelse(min(pull(opf_select, bh)) > 0.70, invisible(">"), invisible("<"))` 0.70).

**Conclusions.** Although our data does not support the hypothesis that short-chain fatty acids decrease in individuals that have colorectal tumors, there may be specific scenarios where short-chain fatty acids may still be beneficial for treatment of colorectal cancer. Alternatively, there may be other metabolites that have not been thoroughly investigated that are more important to human colorectal cancer.

### 2018-05-24 submission/manuscript.tex
add abstract template

>>>>

\textbf{Background.} Certain bacterial community members have been shown
to increase colon tumorigenesis while other taxa have been associated
with normal colons. The some of the taxa associated with normal colons
can use fiber to produce short-chain fatty acids which have been shown
to inhibit tumor growth in model systems. However, the data supporting
the importance of SCFAs in humans is less certain. Here we test the
hypothesis that short-chain fatty acid concentration decreases in
individuals with colorectal tumors.

\textbf{Methods.} We analyzed a cross-sectional (n=490) and pre- and
post-treatment (n=67) group for their concentrations of acetate,
butyrate, and propionate. Analysis also included tumor classification
models using Random Forest, imputed gene relative abundance with
PICRUSt, and metagenomic sequencing on a subset (n=85) of the total
cross-sectional group.

\textbf{Results.} No difference in short-chain fatty acid concentration
was found between normal individuals and patients with adenoma or
carcinoma (P-value \textgreater{} 0.15). There was no difference in the
ability of classification models with or without short-chain fatty acids
added to them in predicting adenoma or carcinoma (P-value \textgreater{}
0.05). Using metagenomic sequencing, there was also no difference in
genes involved with short-chain fatty acid synthesis (P-value
\textgreater{} 0.70).

\textbf{Conclusions.} Although our data does not support the hypothesis
that short-chain fatty acids decrease in individuals that have
colorectal tumors, there may be specific scenarios where short-chain
fatty acids may still be beneficial for treatment of colorectal cancer.
Alternatively, there may be other metabolites that have not been
thoroughly investigated that are more important to human colorectal
cancer.
